### 2014-07-17 ResearchNotes.Rmd
Data loaded in.

Research Notes
========================================================
Just a frame right now.

Load data is from http://www.pjm.com/markets-and-operations/ops-analysis/historical-load-data.aspx
Forecasts are from 
http://www.pjm.com/markets-and-operations/ops-analysis/historical-load-forecasts.aspx

```{r}
GoogleSearch<-read.csv("report.csv")
Loads<-read.csv("Loads.csv")
Forecasts<-read.csv("Forecasts.csv")
```