### 2014-07-17 ResearchNotes.Rmd
Ran at one point.  It gettting stuck on the sql pull to RTO.  Don't know why it can't find the data.frame.

Research Notes
========================================================
Just a frame right now.

Load data is from http://www.pjm.com/markets-and-operations/ops-analysis/historical-load-data.aspx
Forecasts are from 
http://www.pjm.com/markets-and-operations/ops-analysis/historical-load-forecasts.aspx

```{r}
GoogleSearch<-read.csv("report.csv")
Loads<-read.csv("Loads.csv")
Forecasts<-read.csv("Forecasts.csv")
```

Format the Loads data.

```{r}
summary(Loads)
```
Those just look like column numbers.  I want hour 19, X.20 and the date.

```{r}
KeyLoads<-Loads[,c(1, 20)]
names(KeyLoads)<-c("Date","Hour19")
#convert to numeric
KeyLoads$Hour19<-as.numeric(as.character(KeyLoads$Hour19))

#convert to ct Date
KeyLoads$Date<-as.POSIXct(KeyLoads$Date,format='%m/%d/%Y')
summary(KeyLoads)
#clean up loads
rm(Loads)
```

Clean up forecasts

```{r}
KeyForecasts<-Forecasts[,c(1,2,21)]
names(KeyForecasts)<-c("EvalTime","Date","Forecast")

#Fixing time
KeyForecasts$Date<-as.POSIXct(KeyForecasts$Date,format='%m/%d/%Y')
KeyForecasts$EvalTime<-as.POSIXlt(KeyForecasts$EvalTime,format="%m/%d/%Y %H:%M:%S")
summary(KeyForecasts)
#why are there two of each date?  Because they do day ahead and two day ahead.
```

Edit the forecasts so that we have only the earliest day ahead forecast.

```{r}
#Keep only the 5am forecasts
KeyForecasts<-KeyForecasts[as.POSIXlt(KeyForecasts$EvalTime,format="%m/%d/%Y %H:%M:%S")$hour==5,]

#Keep only day ahead
KeyForecasts<-KeyForecasts[(as.POSIXct(KeyForecasts$EvalTime,format="%m/%d/%Y %H:%M:%S") -as.POSIXct(KeyForecasts$Date,format='%m/%d/%Y'))<0,]
summary(KeyForecasts)

```

Now work on the GoogleSearch terms.  This is only for Pennsylvania.

```{r}
summary(GoogleSearch)

```
Date is actually ISO.  Hope that does not cause a problem.

Merging the three sets.  Just to be clear I need the acutal, forecast from day before and search result from two days before.  That is what I am testing, if I can improve the next day forecast with previous day searches.  Lots of fun date matching ahead.

```{r}
summary(KeyForecasts)
summary(KeyLoads)
library(sqldf)

#Join the two parts form the RTO
KeyForecasts$Day<-KeyForecasts$Date
KeyLoads$Day<-KeyLoads$Date

RTO<-sqldf("select KeyForecasts.Day, Forecast, Hour19 from KeyForecasts, KeyLoads where KeyForecasts.Day=KeyLoads.Day;")

GoogleSearch$NextDay<-as.POSIXct(GoogleSearch$Day)+24*60*60

Joint<-sqldf("select RTO.*, GoogleSearch.* from RTO, GoogleSearch where NextDay=Date;")
#fix the forecast again
Joint$Hour19<-as.numeric(as.character(Joint$Hour19))

```
Yes, not a lot of overlap.  Can only do last 30 days with trends for daily.

```{r}
summary(Joint)
summary(lm(Hour19~Forecast+Past_30_days,data=Joint))
```
Looks like it is significant!

```{r}
WO<-lm(Hour19~Forecast,data=Joint)
With<-lm(Hour19~Forecast+Past_30_days,data=Joint)

anova(WO,With)
```

check the change in SE

```{r}
summary(WO)$sigma
summary(With)$sigma

#percent decrease in SE is
(summary(WO)$sigma-summary(With)$sigma)/summary(WO)$sigma
```

Nice! 14% reduction in se.  Proof of concept.