### 2014-07-23 ResearchNotes/GoogleTrendLinks.Rmd
Grabed the trends data for 90 days and restructured directory.

Google Trend Links
========================================================
These will get google trends for most of the stats in the PJM teritory.

http://www.google.com/trends/explore#q=weather&geo=US-DE&date=today%203-m&cmpt=date
http://www.google.com/trends/explore#q=weather&geo=US-KY&date=today%203-m&cmpt=date
http://www.google.com/trends/explore#q=weather&geo=US-MD&date=today%203-m&cmpt=date
http://www.google.com/trends/explore#q=weather&geo=US-NJ&date=today%203-m&cmpt=date
http://www.google.com/trends/explore#q=weather&geo=US-OH&date=today%203-m&cmpt=date
http://www.google.com/trends/explore#q=weather&geo=US-PA&date=today%203-m&cmpt=date
http://www.google.com/trends/explore#q=weather&geo=US-VA&date=today%203-m&cmpt=date
http://www.google.com/trends/explore#q=weather&geo=US-WV&date=today%203-m&cmpt=date