### 2014-07-24 ResearchNotes/ResearchNotes.Rmd
Recreated the proof of concept with hour 19 and checked the other trends ideas.

---
title: "ResearchNotes"
author: "James Woods"
date: "07/24/2014"
output: html_document
---
Read in the data file created by the data manipulation

```{r}
load("~/Research/Forecast-Peak-With-Google-Trends/DataManipulation/WTrends")
```

Replicate the hour 19 experiment from the proof of concept.  First how well do the forecasts work.
```{r}
summary(lm(HE19~F19,data=WTrends))

```
Looks like they are a little high most of the times by 3,400 MW but they comove the right way.

Now to add just PA
```{r}
summary(lm(HE19~F19+PATrends,data=WTrends))

```
Try with something else

```{r}
summary(lm(HE19~F19+NewsTrends,data=WTrends))

```
Oh, that works too

```{r}
summary(lm(HE19~F19+GasTrends,data=WTrends))

```
Gas does not

```{r}
summary(lm(HE19~F19+TrafficTrends,data=WTrends))

```
Traffic does not
```{r}
summary(lm(HE19~F19+RestaurantTrends,data=WTrends))

```
Restaurant does not
```{r}
summary(lm(HE19~F19+MovieTrends,data=WTrends))

```
Snort...Movie does.  Cool place to be.

]