### 2014-07-28 Paper/DraftPaper.Rnw
Added draft paper with outline and blank bib file.

\documentclass{article}

\usepackage{longtable}
\usepackage[margin=1in]{geometry}
\usepackage[section]{placeins}
\usepackage{cite}

\title{Improving Day Ahead Electricity Load Forecasts with Google Trends}
\author{Cameron Mulder\\ James Woods}

% \date{}

\begin{document}
\maketitle
\setkeys{Gin}{width=0.7\textwidth}

\begin{abstract}

\end{abstract}

\SweaveOpts{concordance=TRUE}

\section{Introduction}

<<echo=FALSE,results=hide>>=
load("~/Research/Forecast-Peak-With-Google-Trends/DataManipulation/WTrends")
@

\begin{enumerate}
  \item Intro to short term load forecasting.
  \item Why crowd sourced, non technical,  information could be useful.
  \item Google trends is the summation of Google searches.
  \item Outline of paper
\end{enumerate}

\section{Data Sources}

  \subsection{PJM Load Forecasts and Actuals}

    \begin{enumerate}
      \item Data sources.
      \item Documentation of forecasting.
      \item Forecast bias
      \item Statistically adjusted forecasts.
      \item Note that almost all hours are biased and that co-movements are good for peak hours
    \end{enumerate}

  \subsection{Google Trends}

    \begin{enumerate}
      \item Where to get the data
      \item Limitations
      \item Forming a population weighted index.
      \item Other common searches that will be used as counter examples.
    \end{enumerate}

\section{Post Forecast Addition of Google Trends Data}

  \begin{enumerate}
    \item Simple hourly models with Trends.
    \item Gross comparison with actual forecast and statistically adjusted forecasts.
    \item Why this is insufficient.
  \end{enumerate}

  \subsection{Drop Forward Cross-validation}

    \begin{enumerate}
      \item Cross validation concepts.
      \item Why drop forward cross validation is the right concept.
      \item Comparison of drop forward statistically adjusted and Trends adjusted with gross comparisons.
      \item Reiteration that comparison with raw forecasts is a slam dunk.
    \end{enumerate}

  \subsection{Counter-factual Test with Other Common Google Searches}
    \begin{enumerate}
      \item Comparison with: news, recipe, traffic, gas.
      \item Note that some of them kinda work.
    \end{enumerate}

\section{Summary and Conclusions}

% \bibliography{Trends.bib}
% \bibliographystyle{apalike}

\end{document}