### 2014-09-16 ResearchNotes/GetTrendsData.Rmd

---
title: "GetTrendsData.Rmd"
author: "James Woods"
date: "09/11/2014"
output: html_document
---
Google Trend Links
========================================================
These will get google trends for most of the stats in the PJM teritory.

http://www.google.com/trends/explore#q=weather&geo=US-DE&date=today%203-m&cmpt=date
http://www.google.com/trends/explore#q=weather&geo=US-KY&date=today%203-m&cmpt=date
http://www.google.com/trends/explore#q=weather&geo=US-MD&date=today%203-m&cmpt=date
http://www.google.com/trends/explore#q=weather&geo=US-NJ&date=today%203-m&cmpt=date
http://www.google.com/trends/explore#q=weather&geo=US-OH&date=today%203-m&cmpt=date
http://www.google.com/trends/explore#q=weather&geo=US-PA&date=today%203-m&cmpt=date
http://www.google.com/trends/explore#q=weather&geo=US-VA&date=today%203-m&cmpt=date
http://www.google.com/trends/explore#q=weather&geo=US-WV&date=today%203-m&cmpt=date

http://www.google.com/trends/explore#q=gas&date=today%203-m&cmpt=q
http://www.google.com/trends/explore#q=Movie&date=today%203-m&cmpt=q
http://www.google.com/trends/explore#q=News&date=today%203-m&cmpt=q
http://www.google.com/trends/explore#q=Restaurant&date=today%203-m&cmpt=q
http://www.google.com/trends/explore#q=Traffic&date=today%203-m&cmpt=q