### 2014-10-28 LongDataResearchNotes/ARModel.Rmd
Read data and on the way to figuring out how to do the seven day lag.

Working with the Long Data and heading to an AR model
========================================================

Load in data
```{r}
load("~/Forecast-Peak-With-Google-Trends/LongWtrends.RData")
summary(LongWtrends)
```

Data handling
```{r}
LongWtrends$hour<-as.factor(LongWtrends$hour)

```