### 2014-10-28 LongDataResearchNotes/ARModel.Rmd
Created new dataframe with the lagged values.

Working with the Long Data and heading to an AR model
========================================================

Load in data
```{r}
load("~/Forecast-Peak-With-Google-Trends/LongWtrends.RData")
summary(LongWtrends)
```

Data handling
```{r}
LongWtrends$hour<-as.factor(LongWtrends$hour)

#create a date only variable not date time
LongWtrends$DateOnly<-as.Date(LongWtrends$Date)

library(sqldf)

#used to merge on the lagged values
CopyLong<-LongWtrends

LagFormData<-sqldf("select  LongWtrends.DateOnly, LongWtrends.hour, LongWtrends.F, LongWtrends.HE, CopyLong.F as LagF, CopyLong.HE as LagHE, LongWtrends.Weather, CopyLong.Weather as LagWeather  from LongWtrends,CopyLong where LongWtrends.DateOnly=(CopyLong.DateOnly-7) and LongWtrends.hour=CopyLong.hour;   ")

```