### 2014-10-28 LongDataResearchNotes/ARModel.Rmd
Base AR model in place.  Checking for alternate structures

Working with the Long Data and heading to an AR model
========================================================

Load in data
```{r}
load("~/Forecast-Peak-With-Google-Trends/LongWtrends.RData")
summary(LongWtrends)
```

Data handling
```{r}
LongWtrends$hour<-as.factor(LongWtrends$hour)

#create a date only variable not date time
LongWtrends$DateOnly<-as.Date(LongWtrends$Date)

library(sqldf)

#used to merge on the lagged values
CopyLong<-LongWtrends

LagFormData<-sqldf("select  LongWtrends.DateOnly, LongWtrends.hour, LongWtrends.F, LongWtrends.HE, CopyLong.F as LagF, CopyLong.HE as LagHE, LongWtrends.Weather, CopyLong.Weather as LagWeather  from LongWtrends,CopyLong where LongWtrends.DateOnly=(CopyLong.DateOnly-7) and LongWtrends.hour=CopyLong.hour;   ")

summary(LagFormData)

```

First test regression for hour 19

```{r}
summary(lm(HE~F+LagHE+Weather+LagWeather,data=LagFormData[LagFormData$hour=='19',]   ))
library(lmtest)

dwtest(HE~F+LagHE+Weather+LagWeather,data=LagFormData[LagFormData$hour=='19',])

```

Looks nice.  Lets do AR
```{r}
library(nlme)
summary(gls(HE~F+LagHE+Weather+LagWeather, data=LagFormData[LagFormData$hour=='19',],corAR1(form=~1),na.action=na.omit))
```

So the action goes away.  Lets assume AR structure is with previous hour and not previous day.  This will take a while to run.  There are a lot of parameters.
```{r}
summary(gls(HE~hour+(F+LagHE+Weather+LagWeather):hour, data=LagFormData,corAR1(form=~1),na.action=na.omit))
```

FULL WIN.  It works.