### 2014-10-28 LongDataResearchNotes/ARModel.Rmd
Several itterations of the model but nothing that is satisfactory.  Will continue with specification

Working with the Long Data and heading to an AR model
========================================================

Load in data
```{r}
load("~/Forecast-Peak-With-Google-Trends/LongWtrends.RData")
summary(LongWtrends)
```

Data handling
```{r}
LongWtrends$hour<-as.factor(LongWtrends$hour)

#create a date only variable not date time
LongWtrends$DateOnly<-as.Date(LongWtrends$Date)

library(sqldf)

#used to merge on the lagged values
CopyLong<-LongWtrends

LagFormData<-sqldf("select  LongWtrends.DateOnly, LongWtrends.hour, LongWtrends.F, LongWtrends.HE, CopyLong.F as LagF, CopyLong.HE as LagHE, LongWtrends.Weather, CopyLong.Weather as LagWeather  from LongWtrends,CopyLong where LongWtrends.DateOnly=(CopyLong.DateOnly+7) and LongWtrends.hour=CopyLong.hour;   ")

summary(LagFormData)

```

First test regression for hour 19

```{r}
summary(lm(HE~F+LagHE+Weather+LagWeather,data=LagFormData[LagFormData$hour=='19',]   ))
library(lmtest)

dwtest(HE~F+LagHE+Weather+LagWeather,data=LagFormData[LagFormData$hour=='19',])

```

Looks nice.  Lets do AR
```{r}
library(nlme)
summary(gls(HE~F+LagHE+Weather+LagWeather, data=LagFormData[LagFormData$hour=='19',],corAR1(form=~1),na.action=na.omit))
```

So the action goes away.  Lets assume AR structure is with previous previous day within hour.  This will take a while to run.  There are a lot of parameters.

```{r}
LagFormData<-sqldf("select * from LagFormData order by hour, DateOnly;")

summary(gls(HE~hour+(F+LagHE+Weather+LagWeather):hour, data=LagFormData,corAR1(form=~1),na.action=na.omit))$tTable
```
Sweet works except hours 1-5

Now go by hour within day

```{r}
LagFormData<-sqldf("select * from LagFormData order by DateOnly,hour;")
ARbyHour<-gls(HE~hour+(F+LagHE+Weather+LagWeather):hour, data=LagFormData,corAR1(form=~1),na.action=na.omit)
summary(ARbyHour)$tTable
```

and the results are different.  Now we have to figure out which is right, or perhaps both.  The likelyhood suggests hour within day, i.e., the worse one.

Lets get into the day of week effects and see if it is still there.  There was a Monday and Tuesday problem

```{r}
LagFormData$DOW<-as.factor(weekdays(LagFormData$DateOnly))
LagFormData$DayType<-"MidWeek"
LagFormData$DayType[LagFormData$DOW=="Monday"| LagFormData$DOW=="Tuesday"]<-"EarlyWeek"
LagFormData$DayType[LagFormData$DOW=="Saturday"| LagFormData$DOW=="Sunday"]<-"Weekend"
LagFormData$DayType<-as.factor(LagFormData$DayType)
summary(LagFormData)

LagFormData<-sqldf("select * from LagFormData order by DateOnly,hour;")
ARbyHourDay<-gls(HE~-1+(hour+(F+LagHE+Weather+LagWeather):hour):DayType, data=LagFormData,corAR1(form=~1),na.action=na.omit)
summary(ARbyHourDay)$tTable
# 
# ARbyHourDay<-gls(HE~(hour+(F+LagHE+Weather+LagWeather):hour):weekdays(DateOnly), data=LagFormData,corAR1(form=~1),na.action=na.omit)
# summary(ARbyHourDay)$tTable
```

Try without lag of forecast

```{r}

LagFormData<-sqldf("select * from LagFormData order by DateOnly,hour;")
ARbyHourDayNoLagLoad<-gls(HE~-1+(hour+(F+Weather+LagWeather):hour):DayType, data=LagFormData,corAR1(form=~1),na.action=na.omit)
summary(ARbyHourDayNoLagLoad)$tTable

```

We think that lag of actual is in the forecast.

No lag of actual but with lag forecast

```{r}

LagFormData<-sqldf("select * from LagFormData order by DateOnly,hour;")
ARbyHourDayLagF<-gls(HE~-1+(hour+(F+Weather+LagWeather+LagF):hour):DayType, data=LagFormData,corAR1(form=~1),na.action=na.omit)
summary(ARbyHourDayLagF)$tTable

```