### 2014-12-15 CAData/CAISO/LoadData.Rmd
Loading in an processing data from the CAISO on 7 day ahead and actual ISO loads.

---
title: "PrepCAISOData"
author: "James Woods"
date: "12/15/2014"
output: html_document
---

Read in the data from the CAISO

```{r}
setwd("/home/woodsj/Research/Forecast-Peak-With-Google-Trends/CAData/CAISO")
# Get the files names
files <- list.files(pattern="*.csv")

# First apply read.csv, then rbind
CAISORaw<- do.call("rbind", lapply(files, function(x) read.csv(x, stringsAsFactors = FALSE)))

```
Editing of data types.  Remember that most everything gets imported as a string.

```{r}

CAISORaw$MARKET_RUN_ID<-as.factor(CAISORaw$MARKET_RUN_ID)
CAISORaw$TAC_AREA_NAME<-as.factor(CAISORaw$TAC_AREA_NAME)
CAISORaw$StartTimeGMT<-as.POSIXct(CAISORaw$INTERVALSTARTTIME_GMT)
```
Now trim off all but the full ISO area

```{r}
CAISOWorking<-CAISORaw[CAISORaw$TAC_AREA_NAME=='CA ISO-TAC',]
save(CAISOWorking,file="CAISOWorking.Rdata")
```